---
title: "bslib"
output: 
  html_document:
    code_folding: show
    theme:
      # bg: "#202123"
      # fg: "#B8BCC2"
      primary: "#EA80FC"
      secondary: "#00DAC6"
      base_font:
        google: Prompt
      heading_font:
        google: Proza Libre
      runtime: shiny
date: "`r format(Sys.Date(), '%b %d %Y')`"
editor_options: 
  chunk_output_type: console
---

# Dashboards

A dashboard is created by providing a `title`, `sidebar`, and main area contents
to `page_sidebar()`

```{r setup, include=FALSE}
knitr::opts_chunk$set(
message = FALSE, warning = FALSE
)
```


```{r}
# browseURL("https://rstudio.github.io/bslib/articles/dashboards/index.html")
library(shiny)
library(bslib)
library(ggplot2)

ui <- page_sidebar(
  title = "title"
  ,sidebar = "sidebar"
  ,"main content area"
)

shinyApp(ui, function(input, output){})
```

`sidebar` and main content area can hold UI elements. A good practice is to keep
inputs in the `sidebar` and outputs in the main content area. Additionally, 
consider wrapping outputs in a `card()` and `sidebar` contents in a `sidebar()`
object to add a title or customize styling, positioning, etc.

```{r}
data(penguins, package = "palmerpenguins")

ui <- page_sidebar(
  title = "Penguins dashboard"
  ,sidebar = sidebar(
    title = "Histogram controls",
    varSelectInput(
      "var", "select variable"
      ,dplyr::select_if(penguins, is.numeric)
    )
    ,numericInput("bins", "Number of bins", 30)

  )
      ,card(
      card_header("Histogram"), plotOutput('p')
    )
)

server <- function(input, output){
  output$p <- renderPlot({
    ggplot(penguins) +
      geom_histogram(aes(.data[[input$var]]), bins = input$bins)
  })
}

shinyApp(ui, server)
```

## Layering multiple inputs, pages

To achieve this, you need some reusable outputs and inputs - achieved by a list 
of `cards`, each with a `plotOutput()` container for a numeric variable of interest
and a `color_by` input to control coloring

```{r}
cards <- list(
  card(
    full_screen = TRUE,
    card_header("Bill length"),
    plotOutput("bill_length")
  )
  ,card(
    full_screen = TRUE,
    card_header("Bill depth"),
    plotOutput("bill_depth")
  )
  ,card(
    full_screen = TRUE,
    card_header("Body Mass"),
    plotOutput("body_mass")
  )
)

color_by <- varSelectInput(
  "color_by", "Color by",
  penguins[c("species", "island", "sex")],
  selected = "species"
)
```

Now pass `color_by` directly to `sidebar` and the `cards` to the main content area.
Each card is placed in a new row by default - allowed to grow/shrink to fit the
browser window.

```{r}
ui <- page_sidebar(
  title = "Penguins dashboard",
  sidebar = color_by,
  !!!cards
)

server <- function(input, output){
  gg_plot <- reactive({
    ggplot(penguins) +
      geom_density(aes(fill = !!input$color_by), alpha = 0.2) +
      theme_bw(base_size = 16) + 
      theme(axis.title = element_blank())
  })
  
  output$bill_length <- renderPlot(gg_plot() + aes(bill_length_mm))
  output$bill_depth <- renderPlot(gg_plot() + aes(bill_depth_mm))
  output$body_mass <- renderPlot(gg_plot() + aes(body_mass_g))
}

shinyApp(ui, server)
```

## Layouts

### Multi-column

Multiple columns  are created with `layout_columns()` - widths controlled by 
`col_widths`, if not specified, it divides space evenly among the UI elements
 in a row

```{r}
ui <- page_sidebar(
  title = "Penguins dashboard",
  sidebar = color_by,
  layout_columns(
    cards[[1]], cards[[2]], cards[[3]]
  )
)

shinyApp(ui, server)
```


You can supply a vector of column widths to allocate a given number of columns 
(out of 12) to each element. If elements exceed 12 column mark, they get wrapped
onto a new row. Each row is of equal height, modified by passing `row_heights`
argument.

```{r}
ui <- page_sidebar(
  title = "Penguins dashboard",
  sidebar = color_by,
  layout_columns(
    col_widths = c(4, 8, 12),
    row_heights = c(1,2),
    cards[[1]], cards[[2]], cards[[3]]
  )
)

shinyApp(ui, server)
```


Negative `col_width` imply empty columns

```{r}
ui <- page_sidebar(
  title = "Penguins dashboard",
  sidebar = color_by,
  layout_columns(
    col_widths = c(4, 8, -2, 8, -2),
    cards[[1]], cards[[2]], cards[[3]]
  )
)

shinyApp(ui, server)
```


### Multi-page

Change `page_sidebar()` to `page_navbar()` and use `nav_panel()` to supply the 
title and content for each page. You can also use `nav_spacer()` and `nav_item()`
to control the alignment and UI elements

```{r}
ui <- page_navbar(
  title = "Penguins dashboard",
  sidebar = color_by,
  nav_spacer(),
  nav_panel("Bill Length", cards[[1]]),
  nav_panel("Bill Depth", cards[[2]]),
  nav_panel("Body Mass", cards[[3]]),
  nav_item(tags$a("Posit", href = "https://posit.co"))
)

shinyApp(ui, server)
```


### Multi-panel

You use `navset_*()` components

```{r}
ui <- page_sidebar(
  title = "Penguins dashboard",
  sidebar = color_by,
  navset_card_underline(
    title = "Histogram by species",
    nav_panel("Bill Length", plotOutput("bill_length")),
    nav_panel("Bill Depth", plotOutput("bill_depth")),
    nav_panel("Body Mass", plotOutput("body_mass"))
  )
)

shinyApp(ui, server)
```

### Scrolling vs filling

`page_sidebar()` and `page_navbar()` use a default filling layout - elements 
allowed to grow/shrink as needed to fit a window - which sometimes leads to
elements growing/shrinking beyond what is desirable. You can set `height` on
`card()` to mitigate this


```{r}
ui <- page_sidebar(
  title = "Penguins dashboard",
  sidebar = color_by,
  card(min_height = 200, plotOutput("bill_length")),
  card(height = 200, lorem::ipsum(15))
)

shinyApp(ui, server)
```


### Mobile Layout

Enable filling layout on mobile by setting `fillable_mobile = TRUE`

```{r}
ui <- page_sidebar(
  title = "Penguins dashboard",
  sidebar = color_by,
  fillable_mobile = TRUE,
  card(min_height = 200, plotOutput("bill_length")),
  card(height = 200, lorem::ipsum(15))
)

shinyApp(ui, server)
```


## Components

### Value boxes

A `value_box` is a special kind of card designed for highlighting a value along
with a title and a `showcase` placeholder 
([bsicons](https://icons.getbootstrap.com/) icons)

```{r}
means <- colMeans(
  penguins[c("bill_length_mm", "bill_depth_mm", "body_mass_g")],
  na.rm = TRUE
)

ui <- page_sidebar(
  title = "Penguins dashboard",
  sidebar = color_by,
  layout_columns(
    fill = FALSE,
    value_box(
      title = "Average bill length",
      value = scales::unit_format(unit = "mm")(means[[1]]),
      showcase = bsicons::bs_icon("align-bottom")

    ),
    value_box(
      title = "Average bill depth",
      value = scales::unit_format(unit = "mm")(means[[2]]),
      showcase = bsicons::bs_icon("align-center")
    )
    ,value_box(
      title = "Average body mass",
      value = scales::unit_format(unit = "g", big.mark = ",")(means[[3]]),
      showcase = bsicons::bs_icon("handbag")
    )
  ),
  layout_columns(
    cards[[1]], cards[[2]]
  ),
  cards[[3]]
)

shinyApp(ui, server)
```

